A

### 2015-10-16 paper-draft/paper.tex
add images, updates to paper draft

@@

\section{Distributed compact-finite difference solution}

>>>>

\section{Related work}

@@

\section{Evaluation of compact finite differences}

\begin{figure}[h!]
\begin{center}
\includegraphics[height=200pt]{img/computational-domain.eps}
\end{center}
\caption{Computational domain in 3-D}
\label{fig:computational-domain}
\end{figure}

The computational domain (Figure \ref{fig:computational-domain})
is a structured grid comprised of $nx \times ny \times nz$ grid points.
While evaluating the derivative in, say, the x-direction,
the computational domain is thought of as
a collection of \emph{grid lines}---each consisting of
$nx$ grid points---oriented in the x-direction.
For each grid line, we may assemble the system in
Equation \ref{eqn:compact-tridiagonal-system}.
Solving for the derivatives in the $x$ direction
at each grid point in the domain then entails
solving a tridiagonal system for each of the $ny \times nz$ grid lines.
We note immediately that the left-hand-side (LHS) matrix
is the same for each of the tridiagonal systems,
and only the right-hand-side (RHS) vector is different.

\section{Cyclic reduction}

The main step in evaluating compact-finite differences is the
solution of tridiagonal systems
for each grid line in the computational domain.
The standard approach for solving tridiagonal systems is
the Thomas algorithm.
However, on parallel architectures such as GPUs,
it is advantageous to use other algorithms like
cyclic reduction \cite{Zhang2010FTS}.

\begin{figure}[h!]
\begin{center}
\includegraphics[height=200pt]{img/cyclic-reduction.eps}
\end{center}
\caption{Cyclic reduction -
the values that are solved for at each step of backward reduction
are highlighted}
\label{fig:cyclic-reduction}
\end{figure}

The Cyclic Reduction algorithm consists of two phases:
\emph{forward reduction} and \emph{backward substitution}
(see Figure \ref{fig:cyclic-reduction}).
The forward substitution phase begins with a system of size $n$,
and by expressing every even-indexed equation $i$ as a linear
combination of equations $i$, $i-1$ and $i+1$, reduces it to a
system of size $n/2$.
The process is repeated until a system of 2 equations in 2 unknowns
is left.

For each even-indexed $i$, we define:

\begin{equation*}
k_1 = \frac{a_i}{b_{i-1}},
k_2 = \frac{c_i}{b_{i+1}}
\end{equation*}

Then, we update the values of $a$, $b$, $c$ and $d$ as:

\begin{align} \label{eqn:forward-reduction}
& a^{\prime}_i = -a_{i-1}k_1 & \\
& b^{\prime}_i = b_i - c_{i-1}k_1 - a_{i+1}k_2 & \\
& c^{\prime}_i = -c_{i+1}k_2 & \\
& d^{\prime}_i = d_i - d_{i-1}k_1  - d_{i+1}k_2 &
\end{align}

The 2-by-2 system of equations is solved trivially,
yielding $x_n$ and $x_{n/2}$.
In the backward substitution phase,
every odd-indexed unknown $x_i$ is solved for by
substituting the known values of $x_{i-1}$ and $x_{i+1}$:

\begin{align} \label{eqn:back-substitution}
x_i = \frac{d^{\prime}_i - a^{\prime}_ix_{i-1} - c^{\prime}_ix_{i+1}}{b^{\prime}_i}
\end{align}

At each step, for the last index $i=n$,
the forward reduction step is instead:
\begin{align} \label{eqn:forward-reduction-last}
    & a^{\prime}_n = -a_{n-1}k_1 & \\
    & b^{\prime}_n = b_n - c_{n-1}k_1 & \\
    & d^{\prime}_n = d_n - d_{n-1}k_1 &
\end{align}

And the backward substitution step for $i=1$ is instead:
\begin{align} \label{eqn:back-substitution-first}
x_1 = \frac{d^{\prime}_1 - c^{\prime}_1x_{2}}{b^{\prime}_1}
\end{align}

While cyclic reduction performs far more work per step
compared to the Thomas algorithm,
it is able to solve the system in far fewer steps
on massively parallel architectures.
As pointed out by \cite{Zhang2010FTS},
$n$ parallel processors can solve a tridiagonal system
of size $n$ in $2log(n) - 1$ steps.
The Thomas algorithm, on the other hand,
requires $2n$ steps.

\section{Cyclic reduction for near-Toeplitz systems}

We note from
Equations \ref{eqn:forward-reduction}-\ref{eqn:back-substitution-last}
that each forward reduction step involves 12 operations.
As the LHS matrix is constant,
it is advantageous to pre-compute the coefficients appearing
in the forward reduction step.
For a general tridiagonal matrix,
this requires a substantial amount of storage
(2$n$ for each of the arrays $a$, $b$ and $c$):
$\frac{n}{2} + \frac{n}{4} + \frac{n}{8} + ... = n$
for the reduction steps,
and an additional $n$ for the original coefficients.
\section{Distributed compact-finite difference solution}
